---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

if (bigrquery:::secret_can_decrypt("bigrquery")) {
  json <- bigrquery:::secret_read("bigrquery", "service-token.json")
  bigrquery::bq_auth(path = rawToChar(json))
} else {
  knitr::opts_chunk$set(eval = FALSE)
}
```
# bigrquery

<!-- badges: start -->
[![CRAN Status](https://www.r-pkg.org/badges/version/bigrquery)](https://cran.r-project.org/package=bigrquery)
[![Travis build status](https://travis-ci.org/r-dbi/bigrquery.svg?branch=master)](https://travis-ci.org/r-dbi/bigrquery)
[![Codecov test coverage](https://codecov.io/gh/r-dbi/bigrquery/branch/master/graph/badge.svg)](https://codecov.io/gh/r-dbi/bigrquery?branch=master)
  <!-- badges: end -->

The bigrquery package makes it easy to work with data stored in 
[Google BigQuery](https://developers.google.com/bigquery/) by allowing you to query BigQuery tables and retrieve metadata about your projects, datasets, tables, and jobs. The bigrquery package provides three levels of abstraction on top of BigQuery:

* The low-level API provides thin wrappers over the underlying REST API. All 
  the low-level functions start with `bq_`, and mostly have the form 
  `bq_noun_verb()`. This level of abstraction is most appropriate if you're
  familiar with the REST API and you want do something not supported in the 
  higher-level APIs.
  
* The [DBI interface](http://www.r-dbi.org) wraps the low-level API and
  makes working with BigQuery like working with any other database system.
  This is most convenient layer if you want to execute SQL queries in 
  BigQuery or upload smaller amounts (i.e. <100 MB) of data.

* The [dplyr interface](http://dbplyr.tidyverse.org/) lets you treat BigQuery 
  tables as if they are in-memory data frames. This is the most convenient 
  layer if you don't want to write SQL, but instead want dbplyr to write it 
  for you.

## Installation

The current bigrquery release can be installed from CRAN: 

```R
install.packages("bigrquery")
```

The newest development release can be installed from GitHub:

```R
# install.packages('devtools')
devtools::install_github("r-dbi/bigrquery")
```

## Usage

### Low-level API

```{r}
library(bigrquery)
billing <- bq_test_project() # replace this with your project ID 
sql <- "SELECT year, month, day, weight_pounds FROM `publicdata.samples.natality`"

tb <- bq_project_query(billing, sql)
bq_table_download(tb, max_results = 10)
```

## DBI

```{r, warning = FALSE}
library(DBI)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "publicdata",
  dataset = "samples",
  billing = billing
)
con 

dbListTables(con)

dbGetQuery(con, sql, n = 10)
```

### dplyr

```{r, message = FALSE}
library(dplyr)

natality <- tbl(con, "natality")

natality %>%
  select(year, month, day, weight_pounds) %>% 
  head(10) %>%
  collect()
```

## Important details

### Authentication

When using bigquery interactively, you'll be prompted to [authorize bigrquery](https://developers.google.com/bigquery/authorization) in the browser. Your token will be cached across sessions in `~/.R/gargle/gargle-oauth`, by default. For non-interactive usage, you can download a service account token as a JSON file and put it into force via `bq_auth(path = "/path/to/your/service-account.json")`. bigrquery obtains a token with `gargle::token_fetch()`, which supports even more token flows. For full details, such as how to take advantage of Application Default Credentials or service accounts on GCE VMs, read the gargle vignette [How gargle gets tokens](https://gargle.r-lib.org/articles/how-gargle-gets-tokens.html).

Note that `bigrquery` requests permission to modify your data; but it will never do so unless you explicitly request it (e.g. by calling `bq_table_delete()` or `bq_table_upload()`).

### Billing project

If you just want to play around with the bigquery API, it's easiest to start with the Google's free [sample data](https://developers.google.com/bigquery/docs/sample-tables). You'll still need to create a project, but if you're just playing around, it's unlikely that you'll go over the free limit (1 TB of queries / 10 GB of storage).

To create a project:

1. Open https://console.cloud.google.com/ and create a project.
   Make a note of the "Project ID" in the "Project info" box.

1. Click on "APIs & Services", then "Dashboard" in the left the left menu.

1. Click on "Enable Apis and Services" at the top of the page, 
   then search for "BigQuery API" and "Cloud storage".

Use your project ID as the `billing` project whenever you work with free sample data; and as the `project` when you work with your own data.

## Useful links

* [SQL reference](https://developers.google.com/bigquery/query-reference)
* [API reference](https://developers.google.com/bigquery/docs/reference/v2/)
* [Query/job console](https://bigquery.cloud.google.com/)
* [Billing console](https://console.cloud.google.com/)

## Policies

Please note that the 'bigrquery' project is released with a [Contributor Code of Conduct](https://bigrquery.r-dbi.org/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

[Privacy policy](https://www.tidyverse.org/google_privacy_policy)
